Individual heterogeneity in the functional network topography of the medial prefrontal cortex

### 2019-05-03 Manuscript.Rmd
two discussion points left, plus reinstatement of eigenvalue connectedness info

In order to evaluate the validity of the resulting partitionings across community-detection methods, we also estimated network communities using the more traditional approach of modularity maximization [@Garcia2017], based on the algorithm from Clauset et al. [-@Clauset2004]. This method compares the strength of the connection between any two vertices to their probability of being randomly connected to any other vertex in the network. The method heuristically iterates through many possible combinations of vertices, and selects the partitioning that maximizes the within-community edge weights, relative to a random network containing the same number of edges. Unlike SP, modularity can fractionate a network into more than two communities. Agreement between the partitions provided by the bounded and unbounded community detection methods would suggest the results are not distorted by the restriction of SP to binary partitionings..

@@

@@

The two clustering methods had high agreement (mean ARI = `r round(mean(RI_all_within$Index), digits = 2)`, SD = `r round(sd(RI_all_within$Index), digits = 2)`). Modularity showed a tendency to produce additional communities (median = `r median(nModularity)`, range = `r range(nModularity)`). However, the additional communities encompassed a small number of vertices (median = `r median(modularityCounts_extra)`, IQR = `r quantile(modularityCounts_extra, 0.25)` - `r quantile(modularityCounts_extra, 0.75)`) compared to the principal two (median = `r median(modularityCounts_two)`, IQR = `r quantile(modularityCounts_two, 0.25)` - `r quantile(modularityCounts_two, 0.75)`), suggesting that a binary partitioning could provide a reasonable approximation of the network's true community structure.

@@

Here we address these challenges by using spectral partitioning (SP), a graph-theoretic community detection algorithm that efficiently separates a network into two [@Fiedler1975; @Higham2007; @Tian2018]. For 100 individuals, we reliably subdivided brain regions that typically show overlap between DN and valuation effects into DN and non-DN communities. Restricting our analyses to a general mPFC/PCC search space made it appropriate to use a technique that identified a vertex-wise, binary partitioning that was sensitive to the complex topography of the brain. This contrasts with whole-brain network analyses, which need to allow for multiple sub-networks and which often use coarser-grained parcels as the units of analysis. Partitioning an individual's brain network through SP has a number of advantages, such as identifying communities deterministically, constraining comunities to contain a similar number of vertices (i.e. preventing the allocation of most vertices to a single community), providing continuous values to determine the strength of a node's community affiliation, and diagnosing the connectedness and separability of a network through examination of its resulting eigenvalues (**TODO: show this in text**) [@chung1997; @Tian2018]. Comparisons with partitionings formed by modularity maximization, which heuristically determines the ideal number of communities, suggested the binary partitioning was appropriate.

We found a generalizable pattern across individual partitionings, in which the DN community covered ventral/dorsal mPFC and posterior PCC, with the non-DN community concentrated in pregenual ACC and anterior PCC. However, the precise spatial organization of this general community structure was highly heterogenous across individuals, yet stable across test/re-test evaluations within-individuals. This idiosyncrasy in functional topography was particularly pronounced in mPFC, and was present with various amounts of data.

The degree of individual-specifitcity observed in this work can arise from a variety of sources. For example, the functional topography of mPFC could be governed by its underlying sulcal and gyral organization, which has been shown to systematically vary across individuals [@Mackey2014]. Individual variability could also be due to shifts in functional organization that are independent of structural features [@Haxby2013], or be characterized by the pattern of functional connections with the rest of the brain [@Mars2018; @Passingham2002; @Tobyne2018]. A good goal for future work would be to assess whether the network layout in this region can be predicted on the basis of aspects of brain structure, such as sulcal morphology, myeloarchitecture [@Glasser2016], or structural connectivity [@Saygin2016].

@@

Our results are complementary with previous studies that described individual-specific brain organization. On one hand, topographical heterogeneity has been qulitatively highlighted using a different data aspect ratio [few individuals and more data per individual; @Gordon2017, @Braga2017, @Braga2019]. In terms of stability, previous work has relied on parcellations and group priors to show that an individual's functional connectivity pattern is consistent in time [@Gratton2018; @Kong2018], thus being relatively insensitive to an individual's complex topography. Here we were able to quantify the variability and stability of fine-scale functional topography in a large sample with moderately low amounts of data (down to a single 15 minute scan).

@@

  - We found SP results could be made more stable by thresholding the underlying Fiedler vector. We used temporal stability to apply a custom threshold per subject, but there might be ways to use a generalized threshold in the future.

@@

<!--
- Summarize the results
- Ours are not the first to show variability across individuals, but previous studies have focused on a relatively small number of subjects and often rely on qualitative assessment of these differences. They also use parcellations for networks, whereas we focused on individual vertices to gain spatial finesse. Also, to the best of our knowledge we are the first to quantify the organizational heterogeneity of these areas in the human brain. 
- We are also not the first to show functional subdivision of DN [@Kernbach2018] (but that's using group averages on parcellated data), or this particular individualized pattern of subdivision (braga). Our results are complementary in the sense that we show similar outcomes in 100 subjects. The major discrepancy is that we assign meaning to the communities based on their coverage of regions identified through a meta-analysis.
- Stability within subject has also been shown before. This work was done using either parcellations and oversampling though. [@Laumann2015; @Gordon2017; @Braga2017; @Gratton2018b; @Kong2018]
- Talk about SP as a way to identify an individual's partition: deterministic, with simple diagnostics, and requires less data requirements.
- Similarity with B&B: One way in which our results are at odds with those from Braga and Buckner is the functional labeling of these sub-networks. Based on our criteria, network B would more closely identify with DN (per 7m affiliation), while we currently do not have enough evidence to label network A. These discrepancies should be addressed with explicit behavioral manipulations in the future. Also our search space is limited compared to theirs, thus making the patterns only partially similar.
- The main gap is mapping behavior to functional organization, but we will do that with behavioral manipulations of subjective value in the future. Instead, we focus on understanding the importance of individualized mapping in these areas first, as a baseline.
- Clinical implications: individualized targets, and more accurate estimation of impacts based on assessment of individual patients. Similarly, our approach focuses on individual estimations without group priors, so it's well fit to address the problem of low n in clinical research. Also, SP can produce these effects with moderate amounts of data. Previous results suggest impractical amounts.

>>>>

In order to evaluate the validity of the resulting partitionings across community-detection methods, we also estimated network communities using the more traditional approach of modularity maximization [@Garcia2017], based on the algorithm from Clauset et al. [-@Clauset2004]. This method compares the strength of the connection between any two vertices to their probability of being randomly connected to any other vertex in the network. The method heuristically iterates through many possible combinations of vertices, and selects the partitioning that maximizes the within-community edge weights, relative to a random network containing the same number of edges. Unlike SP, modularity can fractionate a network into more than two communities. Agreement between the partitions provided by the bounded and unbounded community detection methods would suggest the results are not distorted by the restriction of SP to binary partitionings.

<!--
We also diagnosed our choice of two communities by reporting the rank (from lowest to largest) of the first eigenvalue produced by SP to be non-zero. This eigenvalue serves as an estimator of the optimal number of communities in a network (Chung 1999), such that a rank of two in our partitions justifies limiting the number of communities to two.
-->

@@

# # eigenvalue-based diagnosis of optimal n of communities
# prcntDissection <- lapply(Summaries, function(data) {which(data$EigenVal < 0.00001)}) # how many consecutive 0s?
# prcntDissection <- mean((sapply(prcntDissection, length) + 1) == 2) * nSubj # % of subjects who only had 1 zero (thus optimal cuts = 2)

@@

The two clustering methods had high agreement (mean ARI = `r round(mean(RI_all_within$Index), digits = 2)`, SD = `r round(sd(RI_all_within$Index), digits = 2)`). Modularity showed a tendency to produce additional communities (median = `r median(nModularity)`, range = `r range(nModularity)`). However, the additional communities encompassed a small number of vertices (median = `r median(modularityCounts_extra)`, IQR = `r quantile(modularityCounts_extra, 0.25)` - `r quantile(modularityCounts_extra, 0.75)`) compared to the principal two (median = `r median(modularityCounts_two)`, IQR = `r quantile(modularityCounts_two, 0.25)` - `r quantile(modularityCounts_two, 0.75)`), suggesting that a binary partitioning could provide a reasonable approximation of the network's true community structure. <!--  In addition, the estimation of the optimal number of cuts based on the rank of the first non-zero eigenvalue (see "Partition evaluation" for details) showed that 100 percent of the individual networks were best partitioned into two communities. -->

@@

Here we address these challenges by using spectral partitioning (SP), a graph-theoretic community detection algorithm that efficiently separates a network into two [@Fiedler1975; @Higham2007; @Toker2019]. For 100 individuals, we reliably subdivided brain regions that typically show overlap between DN and valuation effects into DN and non-DN communities. Restricting our analyses to a general mPFC/PCC search space made it appropriate to use a technique that identified a vertex-wise, binary partitioning that was sensitive to the complex topography of the brain. This contrasts with whole-brain network analyses, which need to allow for multiple sub-networks and which often use coarser-grained parcels as the units of analysis. Partitioning an individual's brain network through SP has a number of advantages, such as identifying communities deterministically, constraining comunities to contain a similar number of vertices (i.e. preventing the allocation of most vertices to a single community), providing continuous values to determine the strength of a node's community affiliation, and diagnosing the connectedness and separability of a network through examination of its resulting eigenvalues (**TODO: show this in text**) [@chung1997; @Higham2007]. Comparisons with partitionings formed by modularity maximization, which heuristically determines the ideal number of communities, suggested the binary partitioning was appropriate.

We found a generalizable pattern across individual partitionings, in which the DN community covered ventral/dorsal mPFC and posterior PCC, with the non-DN community concentrated in pregenual ACC and anterior PCC. However, the precise spatial organization of this general community structure was highly heterogenous across individuals, yet stable across test/re-test evaluations within-individuals. This idiosyncrasy in functional topography was particularly pronounced in mPFC, and was present with various amounts of data. This degree of individual-specifitcity can arise from a variety of sources. For example, the functional topography of mPFC could be governed by its underlying sulcal and gyral organization, which has been shown to systematically vary across individuals [@Mackey2014]. Individual variability could also be due to shifts in functional organization that are independent of structural features [@Haxby2013], or be characterized by the pattern of functional connections with the rest of the brain [@Mars2018; @Passingham2002; @Tobyne2018]. A good goal for future work would be to assess whether the network layout in this region can be predicted on the basis of aspects of brain structure, such as sulcal morphology, myeloarchitecture [@Glasser2016], or structural connectivity [@Saygin2016].

@@

Our results are complementary with previous studies that described individual-specific brain organization. On one hand, topographical heterogeneity has been qulitatively highlighted using a different data aspect ratio [few individuals and more data per individual; @Gordon2017, @Braga2017, @Braga2019]. In terms of stability, previous work has relied on parcellations and group priors to show that an individual's functional connectivity pattern is consistent in time [@Gratton2018; @Kong2018], thus being relatively insensitive to an individual's complex topography. Here we were able to quantify the variability and stability of fine-scale functional topography in a large sample with moderately low amounts of data (down to a single 15 minute scan). The motivation to sub-divide DN also stems from recent work by Kernbach et al. [-@Kernbach2018], where they identified specialized communication of parcels within DN with the rest of the brain in a large pool of individuals.

We found SP results could be made more stable by thresholding the underlying Fiedler vector. We generated custom thresholds for each individual by interrogating the relationship between Fiedler vector values and the degree of persistent community affiliation of vertices over time. The magnitude of these eigenvector values has been recently used to characterize the continuous connectivity profile of the insula with the rest of the brain, challenging the notion of discrete parcellations of this region [@Tian2018]. We anticipate that these values will be helpful to future neuroscientific studies in providing gradual information not afforded by other clustering algorithms.

@@

@@

<!--
- We are also not the first to show functional subdivision of DN [@Kernbach2018] (but that's using group averages on parcellated data), or this particular individualized pattern of subdivision (braga). Our results are complementary in the sense that we show similar outcomes in 100 subjects. The major discrepancy is that we assign meaning to the communities based on their coverage of regions identified through a meta-analysis.
- Stability within subject has also been shown before. This work was done using either parcellations and oversampling though. [@Laumann2015; @Gordon2017; @Braga2017; @Gratton2018b; @Kong2018]
- Talk about SP as a way to identify an individual's partition: deterministic, with simple diagnostics, and requires less data requirements.
- Similarity with B&B: One way in which our results are at odds with those from Braga and Buckner is the functional labeling of these sub-networks. Based on our criteria, network B would more closely identify with DN (per 7m affiliation), while we currently do not have enough evidence to label network A. These discrepancies should be addressed with explicit behavioral manipulations in the future. Also our search space is limited compared to theirs, thus making the patterns only partially similar.
- The main gap is mapping behavior to functional organization, but we will do that with behavioral manipulations of subjective value in the future. Instead, we focus on understanding the importance of individualized mapping in these areas first, as a baseline.
- Clinical implications: individualized targets, and more accurate estimation of impacts based on assessment of individual patients. Similarly, our approach focuses on individual estimations without group priors, so it's well fit to address the problem of low n in clinical research. Also, SP can produce these effects with moderate amounts of data. Previous results suggest impractical amounts.
